The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates

### 2024-06-11 index.qmd
changed all tables and rendered

```{r libraries}
#libraries
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
theme_set( theme_minimal() ) # ggplot theme
library(broom) # for nice summaries
library(knitr) # 
library(kableExtra) # more tables options. Can cause problems
library(Rmpfr)
# library(reticulate) # incorporate Python
# use_virtualenv('./pythonvenv', required = TRUE) # load pythong venv from path
# library(neurobase)
```

@@

```{r}
rndmean <- function(clm) {
  return(round(mean(clm),2))
}
rndsd <- function(clm) {
  return(round(sd(clm),2))
}
```

@@

::: {#tbl-mri}
```{=latex}
    \begin{tabular*}{\textwidth}{@{}p{2.5cm}llp{2.5cm}p{2.5cm}@{}}
    \toprule
        ~ & T1w & T2w & pcASL & SWI \\ \midrule
        Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated \\
        Phase-encoding direction & Coronal & Sagittal & Axial & Axial \\
        TR (ms) & 7.74 & 2,300 & 4,680 & 30.9 \\
        TE (ms) & 2.97 & 66.29 & 10.55 & 5.24 \\
        Flip angle & 12$^{\circ}$ & 90$^{\circ}$ & 111$^{\circ}$ & 20$^{\circ}$ \\
        FOV (cm) & 20 & 20 & 24 & 25 \\
        Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256 \\
        In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977 \\
        Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 with zero filling (ZIP2) \\
        Number of slices & 126 & 106 & 50 & 92 \\
        Additional parameters & n/a & n/a & 1,450 ms label period; 2,025 ms pulse label; 24 control-label pairs & n/a \\
        Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s \\ \botrule
    \end{tabular*}
    \footnotetext{T1w = T1-weighted; T2w = T2-weighted; pcASL = pseudo-continuous arterial spin labelling; SWI = susceptibility weighted imaging; FSPG = fast spoiled gradient echo; CUBE = General Electric name of sequence, not an acronym; GRE = gradient echo; ZIP2 = through-plane zero filling interpolation}
```
Technical parameters for MR imaging pulse sequences
:::

@@

::: {#tbl-dem}
```{=latex}
\begin{tabular}{@{}ll@{}}
    \toprule
Variable & Subject Data (n=19) \\
  \midrule
Gestational age, weeks (mean $\pm$ SD)                       & `r round(mean(newdata$GA),2)` $\pm$ `r round(sd(newdata$GA),2)`                                                     \\
Corrected gestational age on scan day, weeks (mean $\pm$ SD) & `r round(mean(newdata$CGAonScanDay),2)` $\pm$ `r round(sd(newdata$CGAonScanDay),2)`                                 \\
Number of male neonates (%)                                  & `r sum(newdata$Sex == "M")` (`r round(sum(newdata$Sex == "M")/length(newdata$Sex)*100,2)`)                          \\
Birth weight, g (mean $\pm$ SD)                              & `r round(mean(newdata$BW),2)` $\pm$ `r round(sd(newdata$BW),2)`                                                     \\
Weight on scan day, g (mean $\pm$ SD)                        & `r round(mean(newdata$Weight.on.Scan.Day..kg.*1000),2)` $\pm$ `r round(sd(newdata$Weight.on.Scan.Day..kg.*1000),2)` \\
Days spent in NICU (median, IQR)                             & `r median(newdata$Total_Days_NICU)`, `r IQR(newdata$Total_Days_NICU)`                                               \\
Days on ventilation (median, IQR)                            & `r median(newdata$Total_Days_Ventilation)`, `r IQR(newdata$Total_Days_Ventilation)`                                 \\
\botrule
\end{tabular}
\footnotetext{SD = standard deviation; IQR = inter quartile range}
```
Demographic and clinical characteristic of the study sample.
:::

@@

::: {#tbl-chistats}
```{=latex}
    \begin{tabular}{@{}lllp{2cm}ll@{}}
    \toprule
        Region & Measure & QSM & Paramagnetic map & p-value & 95\% CI \\ \midrule
        \multirow{ 2}{*}{SSS} & Chi (ppm)     & `r round(mean(newdata$Gavin_SSS_Chi),2)` \pm `r round(sd(newdata$Gavin_SSS_Chi),2)`                     & `r round(mean(newdata$Chisep_SSS_Chi),2)` \pm `r round(sd(newdata$Chisep_SSS_Chi),2)`                     & `r signif(SSSChi.t$p.value,3)`                        & `r round(SSSChi.t$conf.low,2)`, `r round(SSSChi.t$conf.high,2)`           \\
                              & SvO$_{2}$(\%) & `r round(mean(newdata$Gavin_SSSVein_CSvO2*100),2)` \pm `r round(sd(newdata$Gavin_SSSVein_CSvO2*100),2)` & `r round(mean(newdata$Chisep_SSSVein_CSvO2*100),2)` \pm `r round(sd(newdata$Chisep_SSSVein_CSvO2*100),2)` & `r signif(SSSSvO2.t$p.value,3)` & `r round(SSSSvO2.t$conf.low*100,2)`, `r round(SSSSvO2.t$conf.high*100,2)` \\
        \multirow{ 2}{*}{CCV} & Chi (ppm)     & `r round(mean(newdata$Gavin_IntVein_Chi),2)` \pm `r round(sd(newdata$Gavin_IntVein_Chi),2)`             & `r round(mean(newdata$Chisep_IntVein_Chi),2)` \pm `r round(sd(newdata$Chisep_IntVein_Chi),2)`             & `r signif(CCVChi.t$p.value,3)`                        & `r round(CCVChi.t$conf.low,2)`, `r round(CCVChi.t$conf.high,2)`           \\
                              & SvO$_{2}$(\%) & `r round(mean(newdata$Gavin_IntVein_CSvO2*100),2)` \pm `r round(sd(newdata$Gavin_IntVein_CSvO2*100),2)` & `r round(mean(newdata$Chisep_IntVein_CSvO2*100),2)` \pm `r round(sd(newdata$Chisep_IntVein_CSvO2*100),2)` & `r signif(CCVSvO2.t$p.value,3)` & `r round(CCVSvO2.t$conf.low*100,2)`, `r round(CCVSvO2.t$conf.high*100,2)` \\
        \botrule
    \end{tabular}
    \footnotetext{QSM = quantitative susceptibility mapping; CI = confidence interval; SSS = superior sagitall sinus; CCV = central cerebral vein}
```
Summary of acquired physiological parameters. Mean ± SD is shown for chi and SvO~2~ values. The P-value and 95% confidence interval (CI) were obtained through the comparison of values between QSM and paramagnetic maps; (n=19).
:::

>>>>

```{r libraries}
#libraries
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
theme_set( theme_minimal() ) # ggplot theme
library(broom) # for nice summaries
library(knitr) # 
library(kableExtra) # more tables options. Can cause problems
library(Rmpfr)
library(gt)
# library(reticulate) # incorporate Python
# use_virtualenv('./pythonvenv', required = TRUE) # load pythong venv from path
# library(neurobase)
```

@@

```{r}
# function to make rounding means and sd easier
rndmean <- function(clm) {
  return(round(mean(clm),2))
}
rndsd <- function(clm) {
  return(round(sd(clm),2))
}
```

@@

```{r}
#| label: tbl-mri
#| tbl-cap: "Technical parameters for MR imaging pulse sequences"
df <- data.frame(Scan = character(), T1w = character(), T2w = character(), pcASL = character(), SWI = character(), stringsAsFactors = FALSE)
df[1,] <- c("Sequence", "3D FSPGR", "3D CUBE", "Multi-shot 3D fast spin-echo", "3D spoiled GRE flow-compensated")
df[2,] <- c("Phase-encoding direction", "Coronal", "Sagittal", "Axial", "Axial")
df[3,] <- c("TR (ms)", "7.74", "2,300", "4,680", "30.9")
df[4,] <- c("TE (ms)", "2.97", "66.29", "10.55", "5.24")
df[5,] <- c("Flip angle", "12\U00B0", "90\U00B0", "111\U00B0", "20\U00B0")
df[6,] <- c("FOV (cm)", "20", "20", "24", "25")
df[7,] <- c("Acquisition matrix", "512 x 512", "256 x 256", "128 x 128", "256 x 256")
df[8,] <- c("In-plane resolution (mm)", "0.39 x 0.39", "0.78 x 0.78", "1.875 x 1.875", "0.977 x 0.977")
df[9,] <- c("Slice thickness (mm)", "1", "1", "4", "2, reconstructed to 1 with zero filling (ZIP2)")
df[10,] <- c("Number of slices", "126", "106", "50", "92")
df[11,] <- c("Additional parameters", "n/a", "n/a", "1,450 ms label period;\n 2,025 ms pulse label;\n 24 control-label pairs", "n/a")
df[12,] <- c("Scan duration", "4 min 39 s", "5 min 1 s", "5 min 26 s", "5 min 29 s")

footnotetext="T1w = T1-weighted; T2w = T2-weighted; pcASL = pseudo-continuous arterial spin labelling; SWI = susceptibility weighted imaging; FSPG = fast spoiled gradient echo; CUBE = General Electric name of sequence, not an acronym; GRE = gradient echo; ZIP2 = through-plane zero filling interpolation"

if (knitr::is_latex_output()) {
colnames(df)[1] <- ""
df[11,] <- linebreak(df[11,])
df |>
  kbl(format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    align = "l",
    escape = FALSE) |>
    kable_styling(font_size = 8, position = "center", latex_options = c("hold_position", "scale_down"), full_width = FALSE) |>
    footnote(general_title = "", 
           footnote_as_chunk = TRUE,
           threeparttable = TRUE,
           general = footnotetext) |>
    column_spec(1, width="8em") |>
    column_spec(4, width="9em") |>
    column_spec(5, width="9em")
} else {
df |> 
    mutate(across(everything(), ~ str_replace_all(., "\n", "<br>"))) |> 
    gt(rowname_col = "Scan") |> tab_footnote(footnotetext) |>
    fmt_markdown(columns = TRUE) |>
    tab_options(quarto.disable_processing = TRUE)
}
```

@@

```{r}
#| label: tbl-dem
#| tbl-cap: Demographic and clinical characteristic of the study sample.

df <- data.frame(Variable = character(), "Subject" = character(), stringsAsFactors = FALSE)
df[1,] <- c("Gestational age, weeks (mean \U00B1 SD)", paste0(rndmean(newdata$GA), " \U00B1 ", rndsd(newdata$GA)))
df[2,] <- c("Corrected gestational age on scan day, weeks (mean \U00B1 SD) ", paste0(rndmean(newdata$CGAonScanDay), " \U00B1 ", rndsd(newdata$CGAonScanDay)))
df[3,] <- c("Number of male neonates (\\%)", paste0(sum(newdata$Sex == "M"), " (", (round(sum(newdata$Sex == "M")/length(newdata$Sex)*100,2)), ")"))
df[4,] <- c("Birth weight, g (mean \U00B1 SD)", paste0(rndmean(newdata$BW), " \U00B1 ", rndsd(newdata$BW)))
df[5,] <- c("Weight on scan day, g (mean \U00B1 SD)", paste0(rndmean(newdata$Weight.on.Scan.Day..kg.*1000), " \U00B1 ", rndsd(newdata$Weight.on.Scan.Day..kg.*1000)))
df[6,] <- c("Days spent in NICU (median, IQR)", paste0(median(newdata$Total_Days_NICU), ", ", IQR(newdata$Total_Days_NICU)))
df[7,] <- c("Days on ventilation (median, IQR)", paste0(median(newdata$Total_Days_Ventilation), ", ", IQR(newdata$Total_Days_Ventilation)))
df <- df |> rename("Subject data (n = 19)" = Subject)
footnotetext="SD = standard deviation; IQR = inter quartile range"

if (knitr::is_latex_output()) {
df |>
  kbl(format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    align = "lc",
    escape = FALSE) |>
    kable_styling(font_size = 9, position = "center", latex_options = c("hold_position", "scale_down"), full_width = FALSE) |>
    footnote(general_title = "", 
           footnote_as_chunk = TRUE,
           threeparttable = TRUE,
           general = footnotetext)
} else {
df |> 
    mutate(across(everything(), ~ str_replace_all(., "\n", "<br>"))) |> 
    gt() |> 
    cols_align(align = "center", columns = c("Subject data (n = 19)")) |>
    tab_footnote(footnotetext) |>
    fmt_markdown(columns = TRUE) |>
    tab_options(quarto.disable_processing = TRUE)
}
```

@@

```{r}
#| label: tbl-chistats
#| tbl-cap: Summary of acquired physiological parameters. Mean $\pm$ SD is shown for chi and SvO~2~ values. The P-value and 95% confidence interval (CI) were obtained through the comparison of values between QSM and paramagnetic maps; (n=19).

df <- data.frame(Region = character(), Measure = character(), QSM = character(), pmap = character(), pvalue = character(), CI = character(), stringsAsFactors = FALSE)

df[1,] <- c("SSS", "Chi (ppm)", paste0(rndmean(newdata$Gavin_SSS_Chi), " \U00B1 ", rndsd(newdata$Gavin_SSS_Chi)), paste0(rndmean(newdata$Chisep_SSS_Chi), " \U00B1 ", rndsd(newdata$Chisep_SSS_Chi)), signif(SSSChi.t$p.value,3), paste0(round(SSSChi.t$conf.low,2), ", ", round(SSSChi.t$conf.high,2)))
df[2,] <- c("SSS", "SvO\U2082 (\\%)", paste0(rndmean(newdata$Gavin_SSSVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_SSSVein_CSvO2*100)), paste0(rndmean(newdata$Chisep_SSSVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_SSSVein_CSvO2*100)), signif(SSSSvO2.t$p.value,3), paste0(round(SSSSvO2.t$conf.low,2), ", ", round(SSSSvO2.t$conf.high,2)))
df[3,] <- c("CCV", "Chi (ppm)", paste0(rndmean(newdata$Gavin_IntVein_Chi), " \U00B1 ", rndsd(newdata$Gavin_IntVein_Chi)), paste0(rndmean(newdata$Chisep_IntVein_Chi), " \U00B1 ", rndsd(newdata$Chisep_IntVein_Chi)), signif(CCVChi.t$p.value,3), paste0(round(CCVChi.t$conf.low,2), ", ", round(CCVChi.t$conf.high,2)))
df[4,] <- c("CCV", "SvO\U2082 (\\%)", paste0(rndmean(newdata$Gavin_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_IntVein_CSvO2*100)), paste0(rndmean(newdata$Chisep_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_IntVein_CSvO2*100)), signif(CCVSvO2.t$p.value,3), paste0(round(CCVSvO2.t$conf.low,2), ", ", round(CCVSvO2.t$conf.high,2)))

df <- df |> rename("Paramagnetic map" = pmap, "p-value" = pvalue, "95\U0025 CI" = CI)

footnotetext="QSM = quantitative susceptibility mapping; CI = confidence interval; SSS = superior sagitall sinus; CCV = central cerebral vein"

if (knitr::is_latex_output()) {
df |>
  kbl(format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    align = "llcccc",
    escape = FALSE) |>
    kable_styling(font_size = 9, position = "center", latex_options = c("hold_position", "scale_down"), full_width = FALSE) |>
    footnote(general_title = "", 
           footnote_as_chunk = TRUE,
           threeparttable = TRUE,
           general = footnotetext)
} else {
df |> 
    mutate(across(everything(), ~ str_replace_all(., "\n", "<br>"))) |> 
    gt() |> 
    cols_align(align = "center", columns = c("QSM", "Paramagnetic map", "p-value", "95\U0025 CI")) |>
    tab_footnote(footnotetext) |>
    fmt_markdown(columns = TRUE) |>
    tab_options(quarto.disable_processing = TRUE)
}
```
